---
title: "QSURE R Training 2022: Lesson 1"
author: "Karissa Whiting & Mike Curry"
date: "June 7th, 2022"
output:
  xaringan::moon_reader:
    css: [xaringan-themer.css, custom.css]
    lib_dir: libs
    seal: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
---


```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)

set.seed(78987)
library(xaringanthemer)
library(gtsummary)
library(dplyr)
style_mono_accent(
  base_color = "#007CBA", 
  code_inline_color = "#FF5733",
  code_font_size = "0.7em"
)

#xaringan::inf_mr() 
options(htmltools.dir.version = FALSE) 

```

```{r xaringanExtra-freezeframe, echo=FALSE}
xaringanExtra::use_freezeframe()
```


class: inverse, center, title-slide, middle
# QSURE R Training 2022
## `Lesson 2`

### Karissa Whiting & Mike Curry

#### June 9th, 2022

---

## Model Statistical analyses 

.xxlarge[
Will cover:
- linear model
- logistic model
- survival analyses 
]

---

## General modeling formula 

.xxlarge[
- in general `~` is used to separate your outcome on the left hand side and your predictors on the right hand side
- your outcome will always be on the left side of the `~`
- only some univariate tests like `chisq.test()` do not use the `~` notation 
- general  notation: model(outcome ~ covariates, data)
]
- the `stats` package is already loaded in R which will make it easier to use common statisitcal tests

---

## Example of linear model

.xxlarge[
- Continuous outcome 
- `mtcars$vs  <- as.character(mtcars$vs)`
- `mtcars$cyl   <- as.character(mtcars$cyl)`
- `lm(mpg ~ vs + cyl, data = mtcars)`
- Specifying interactions 
- `lm(mpg ~ vs * cyl, data = mtcars)`
- `lm(mpg ~ vs + cyl + vs:cyl, data = mtcars)`


```{r, eval= TRUE}

mtcars$vs  <- as.character(mtcars$vs)
mtcars$cyl   <- as.character(mtcars$cyl)
mod1 <- lm(mpg ~ vs * cyl, data = mtcars)
class(mod1) # class of lm which is a list

names(mod1)

```

]

---

## Example cont.

```{r}

summary(mod1)

```

- while it is nice to see the summary results, you wouldn't present them in this fashion

---

## Example cont. 

.xxlarge[

- `broom` and `gt`/`gtsummary` will help
- `broom` is a package that helps tidy model results into data.frames
- this helps with reporting and you can further format the data.frame and present with `gt`

```{r, message=FALSE}

moddf <- broom::tidy(mod1) %>% #didn't load broom just called one function 
          mutate(p.value = round(p.value,3)) %>% 
          select(-std.error)

gt::gt(moddf)

```

]

---

### gtsummary

.xxlarge[

- `gtsummary` makes it easier to report model and descriptive statistics (more on this in the example)
- For more helpful examples: https://www.danieldsjoberg.com/gtsummary/index.html 
]

---

### gtsummary

```{r}
tbl_regression(mod1)
```

---

### Customizing gtsummary

```{r}

library(labelled)

var_label(mtcars$cyl) <- "Cylinder"
newmodsum <- lm(mpg ~ vs * cyl, data = mtcars) %>% 
tbl_regression() %>% 
    bold_labels() %>% 
    modify_caption("New title for model")


```

---

### Customizing gtsummary

```{r}

newmodsum

```


---

## Logistic models

.xxlarge[

- binary outcome (0/1)
- R will model the '1' as the event by default make sure your variable is coded correctly

```{r}
mtcars$vs <- as.numeric(mtcars$vs)
mtcars$am <- as.character(mtcars$am)

mod2 <- glm(vs ~ am , data = mtcars, family =  "binomial")



```

]

---

## Summarize logistic model

```{r}

tbl_regression(mod2, exponentiate = TRUE) %>% 
    bold_labels()

```


---

## Survival analysis

.xxlarge[
- Outcome is both time and an event (e.g death, progression)
- Have to specify both in model  
- Former MSK employee Emily Zabor put together great materials for survival analysis here: https://www.emilyzabor.com/tutorials/survival_analysis_in_r_tutorial.html


]


---

## Survival analysis

```{r}
library(survival)

lung <- lung %>% 
        mutate(ph.ecog = as.character(ph.ecog),
               sex = as.character(sex))

mod3 <- coxph(Surv(time, status)~ph.ecog+sex,data = lung)
mod4 <- survfit(Surv(time, status) ~ sex,data = lung)



```

---

## Survival analysis

```{r}

tbl_regression(mod3, exponentiate = TRUE)

```

---

### Survival analysis


```{r}

mod4

```